Chemically Selective Transport in a Cross-linked H\textsubscript{II} Phase Lyotropic Liquid Crystal Membrane

### 2019-06-25 Ben_Manuscripts/transport/Final_Draft.tex
Final revision

  %BJC4: not sure if I should add a thesis to this paragraph, reattach above, or just
  % leave as is since it's informational.
  Figure~\ref{fig:membrane_rendering} shows an atomistic rendering of the
  membrane in which we placed solutes. We have listed the solutes studied along
  with the abbreviations that we will use in the charts for the remainder of this
  paper in Table~\ref{table:abbreviations}. Note that, in order limit our
  studies to uncharged species, we have only studied acetic acid in its
  protonated form. The pKa of gallic acid is 4.40 while the pKa of acetic acid is
  4.75.~\cite{chuysinuan_gallic_2009,perrin_pka_1981} Assuming that the pKa of
  gallic acid is similar to NaGA3C11, it is possible for our sodium salted
  monomer to exist at the same time as protonated acetic acid. In any case, we
  are more interested in mechanisms of interaction with specific chemical
  functionality more than the behavior of specific molecules.

  \begin{figure}
  \centering
  \begin{subfigure}{0.325\textwidth}
  \vspace{1.4cm}
  \includegraphics[width=\textwidth]{NaGA3C11.pdf}
  \vspace{1cm}
  \caption{}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{ortho_view.pdf}
  \caption{}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{top_view.pdf}
  \caption{}
  \end{subfigure}
  \caption{The H\textsubscript{II} phase formed by LLC monomer NaGA3C11 (a)
  consists of pores that are approximately straight (b) and packed hexagonally (c).
  In both figures, water molecules that occupy the pores are colored blue, 
  the hydrophilic head groups of the LLC monomers are colored orange and all else is colored gray. 
  The chemical environment within the nanopores is largely determined by the functionality
  of the head groups. Additionally, there is not a well-defined transition
  point from the aqueous pore region to the alkane-dense, hydrophobic tail region,
  generating further complexity.
  }\label{fig:membrane_rendering}
  \end{figure}

@@

  We parameterized the interaction potential for the monomer and solutes using
  the Generalized AMBER Force Field (GAFF)~\cite{wang_development_2004} with the
  Antechamber package \cite{wang_automatic_2006} shipped with
  AmberTools16~\cite{case_ambertools16_2016}.  We use GAFF as it has been used
  extensively with organic molecules.  We assigned atomic charges using the
  am1bccsym method of \texttt{molcharge} included with QUACPAC from OpenEye
  Scientific Software.

  We simulated water molecules using the TIP3P water model because it because
  it is the water model that GAFF and other AMBER family parameters were optmized 
  with.~\cite{florova_explicit_2010} %~\cite{mahoney_five-site_2000}

@@

  We observed water enters into the distal tail region of our system. We define the 
  distal tail region to be ca. 1.5 nm from the pore center based on the minimum in 
  the radial distribution function of water (see Section~\ref{method:rdfs}). Therefore
  we aimed to build our initial configurations with water in both regions, close to
  the expected equilibrium partition. The final amount of water present in the distal
  tail region may or may not be experimentally consistent but it is necessary for our
  results to be thermodynamically consistent; otherwise, water would slowly leak 
  from the pore into the tail region over several hundred nanoseconds.

  To determine the approximate ratio of water in the pores to tails, we equilibrated 4 systems
  where we initially placed water in the pores and in the tails in addition to a water
  reservoir between periodic images. We filled the pores with water by running the 
  GROMACS command \texttt{gmx solvate} on our initial configuration followed by the 
  removal of any water molecules placed in the distal tail region. This method requires
  that we vary the monomer starting radius since the vacant space dictates the pore's 
  water content. We then placed a defined number of water molecules into the distal tail
  region one at a time with a short energy minimization between each insertion.
  We chose to test systems with initial pore radii of 5, 6, 7 and 8 \AA~with tail and 
  pore water compositions given in Table~\ref{S-table:water_content}. We found that 
  systems started with more water in the pores than tails tended to equilibrate relatively
  quickly with an approximate ratio of 3 water molecules in the pores to 2 in the tails.

  We removed the water reservoir and allowed the pore and tail water
  contents to equilibrate with 5 and 10 wt\% total water with a 3:2 ratio
  of pore to tail water. We considered the water content
  equilibrated once the water contents plateaued. The 5 wt\% system did
  not plateau until $\sim$ 600 ns 
  while the 10 wt \% water system equilibrated within the first 100 ns of 
  simulation (see Figure~\ref{S-fig:5wt_offset_equil} and Figure~\ref{S-fig:10wt_offset_equil} of
  the Supporting Information). The equilibrated pores contain 72 \% and 69 \% of the total water in the
  5 and 10 wt \% systems respectively. We cross-linked the equilibrated 
  solvated systems, then allowed them to equilibrate further for 100 ns. 
  The water contents in each region does not change significantly in either case 
  (see Figure~\ref{S-fig:xlinked_solvation_equilibration} of the Supporting Information).
  The `wet' equilibration and cross-linking procedures follow
  the methods outlined in our previous work.

@@

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{total_density_5wt.pdf}
  \caption{5 wt \% water, spatial density of all atoms}\label{fig:total_density_5wt}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{total_water_density_5wt.pdf}
  \caption{5 wt \% water, spatial density of water}\label{fig:total_water_density_5wt}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{total_density_10wt.pdf}
  \caption{10 wt \% water, spatial density of all atoms}\label{fig:total_density_10wt}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \vspace{-0.5cm}
  \includegraphics[width=\textwidth]{total_water_density_10wt.pdf}
  \caption{10 wt \% water, spatial density of water}\label{fig:total_water_density_10wt}
  \end{subfigure}
  \caption{Looking down at the $xy$ plane of the cross-linked LLC membranes, the lowest
  density regions of both membranes are in the pore center and in the distal tail region,
  furthest from the pore center (a \& c). Water concentration is highest in regions of
  low monomer density regions (b \& d).}\label{fig:total_density}
  \end{figure}

@@

  We observed transport of sodium, water and 20 other small polar solutes
  inside the membrane nanopores. First, we comment on transport of the
  membrane constituents, water and sodium, in a system absent of any additional
  solutes. We then present the general trends that we observe among the
  all other solutes studied.

@@

  % Changing does not change conclusion here.
  In the 5 wt\% water system, water molecules that spend the majority of their 
  time in the distal tail region have MSDs slightly higher than those of water molecules
  close to the pore center. The MSD of distal tail water molecules is 1.7 nm$^2$
  compared to 1.2 nm$^2$ outside the distal tails. This behavior is likely due to the
  much more crowded pore environment in the 5 wt\% versus the 10 wt\% water system. 
  Monomer head groups are concentrated about 2\AA~closer to the pore centers in the
  5 wt\% versus the 10 wt\% water system (see Figure~\ref{fig:component_densities})  %BJC: this is the RDF figure of head groups, tails, sodium, water 
  which leaves less space for water molecules and restricts their movement among the 
  relatively immobile monomer head groups. The consequently high energetic barrier 
  to diffusion enhances the favorability of associative interactions between water
  and monomer components. The 95th percentile lifetime of hydrogen bonds between 
  water molecules located near the pore center is 2.1 ns in the 5 wt\% system compared
  to 0.5 ns in the 10 wt\% system. However, since we only saved simulation output 
  every 0.5 ns, it is likely that the true hydrogen bond lifetimes in the 10 wt\% 
  system are shorter than this value. Additionally, there are more frequent and longer
  periods of association between water and sodium ions. An average of 65\% of water 
  molecules in the pores of the 5 wt\% water system associate with a sodium ion each
  frame compared to 44\% in the 10 wt\% water system.There are about half as many water
  molecules in the 5 wt\% water system which allows a larger fraction of them to 
  associate with sodium ions. The lifetime of associations between sodium ions and 
  water is 7.5 ns and 3.0 ns in 5 and 10 wt\% water systems respectively.

@@

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\linewidth]{msd_radius_diols_10wt.pdf}
  \caption{}\label{fig:polyols_msd}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\linewidth]{polyols_rdf.pdf}
  \caption{}\label{fig:polyols_rdf}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \vspace{-0.275cm}
  \includegraphics[width=\linewidth]{multi_hbonds.pdf}
  \caption{}\label{fig:multi_hbonds}
  \end{subfigure}
  \caption{(a) The MSDs of diols, triols and sugars descend in order of their size, except
  for propylene glycol which moves exceptionally slow. The MSD of ethylene glycol 
  is in close agreement with the theoretical lines, implying that the solute 
  is subject to a similar amount of hindrance as methanol, the solute to which the
  theoretical lines were fit. (b) Glycerol, tetrose and ribose have high densities 
  close to the pore center because they have a high number of hydrophilic groups and
  are relatively large. Ethylene glycol and propylene glycol are densest close to the
  head group region. (c) The number of hydrogen bond interactions between solutes and
  monomers increases as solutes gain additional hydroxyl groups. The number of 
  hydrogen bonds made by a single solute in different locations simultaneously, n, 
  also increases with the number of hydroxyl groups. In the most extreme case, all 
  four hydroxyl groups of Ribose (n = 4) are involved in a hydrogen bond interaction
  at the same time.}\label{fig:polyols}
  \end{figure}

>>>>

  %BJC4: not sure if I should add a thesis to this paragraph, reattach above, or just
  % leave as is since it's informational.
  Figure~\ref{fig:membrane_rendering} shows an atomistic rendering of the
  membrane in which we placed solutes. We have listed the solutes studied along
  with the abbreviations that we will use in the charts for the remainder of this
  paper in Table~\ref{table:abbreviations}. Note that, in order limit our
  studies to uncharged species, we have only studied acetic acid in its
  protonated form. The pKa of gallic acid is 4.40 while the pKa of acetic acid is
  4.75.~\cite{chuysinuan_gallic_2009,perrin_pka_1981} Assuming that the pKa of
  gallic acid is similar to Na-GA3C11, it is possible for our sodium salted
  monomer to exist at the same time as protonated acetic acid. In any case, we
  are more interested in mechanisms of interaction with specific chemical
  functionality more than the behavior of specific molecules.

  \begin{figure}
  \centering
  \begin{subfigure}{0.325\textwidth}
  \vspace{1.4cm}
  \includegraphics[width=\textwidth]{NaGA3C11.pdf}
  \vspace{1cm}
  \caption{}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{ortho_view.pdf}
  \caption{}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{top_view.pdf}
  \caption{}
  \end{subfigure}
  \caption{The H\textsubscript{II} phase formed by LLC monomer Na-GA3C11 (a)
  consists of pores that are approximately straight (b) and packed hexagonally (c).
  In both (b) and (c), water molecules that occupy the pores are colored blue, 
  the hydrophilic head groups of the LLC monomers are colored orange and all else is colored gray. 
  The chemical environment within the nanopores is largely determined by the functionality
  of the head groups. Additionally, there is not a well-defined transition
  point from the aqueous pore region to the alkane-dense, hydrophobic tail region,
  generating further complexity.
  }\label{fig:membrane_rendering}
  \end{figure}

@@

  We parameterized the interaction potential for the monomer and solutes using
  the General AMBER Force Field (GAFF)~\cite{wang_development_2004} with the
  Antechamber package \cite{wang_automatic_2006} shipped with
  AmberTools16~\cite{case_ambertools16_2016}. We use GAFF as it has been used
  extensively with organic molecules. We assigned atomic charges using the
  am1bccsym method of \texttt{molcharge} included with QUACPAC from OpenEye
  Scientific Software.

  We simulated water molecules using the TIP3P water model because it is the
  water model that GAFF and other AMBER family parameters have been optmized
  with.~\cite{florova_explicit_2010} %~\cite{mahoney_five-site_2000}

@@

  We observed that water enters into the distal tail region of our system. We
  define the distal tail region to be ca. 1.5 nm from the pore center based on
  the minimum in the radial distribution function of water (see
  Section~\ref{method:rdfs}). Therefore we aimed to build our initial
  configurations with water in both regions, close to the expected equilibrium
  partition. The final amount of water present in the distal tail region may or
  may not be experimentally consistent but it is necessary for our results to be
  thermodynamically consistent; otherwise, water would slowly leak from the pore
  into the tail region over several hundred nanoseconds.

  To determine the approximate ratio of water in the pores to tails, we
  equilibrated 4 systems where we initially placed water in the pores and in the
  tails in addition to a water reservoir between periodic images. We filled the
  pores of our initial configuraiton with water by running the GROMACS command
  \texttt{gmx solvate} on our initial configuration followed by the removal of
  any water molecules placed in the distal tail region. This method requires that
  we vary the monomer starting radius since the vacant space dictates the pore's
  water content. We then placed a defined number of water molecules into the
  distal tail region one at a time with a short energy minimization between each
  insertion. We chose to test systems with initial pore radii of 5, 6, 7 and 8
  \AA~with tail and pore water compositions given in
  Table~\ref{S-table:water_content}. We found that systems started with more
  water in the pores than tails tended to equilibrate relatively quickly with an
  approximate ratio of 3 water molecules in the pores to 2 in the tails.

  We removed the water reservoir and allowed the pore and tail water contents
  to equilibrate with 5 and 10 wt\% total water with a 3:2 ratio of pore to tail
  water. We considered the water content equilibrated once the number of water
  molecules in each region plateaued. The 5 wt\% system did not plateau until
  $\sim$ 600 ns while the 10 wt \% water system equilibrated within the first 100
  ns of simulation (see Figure~\ref{S-fig:5wt_offset_equil} and
  Figure~\ref{S-fig:10wt_offset_equil} of the Supporting Information). The
  equilibrated pores contain 72 \% and 69 \% of the total water in the 5 and 10
  wt \% systems respectively. We cross-linked the equilibrated solvated systems,
  then allowed them to equilibrate further for 100 ns. The water contents in
  each region does not change significantly in either case (see
  Figure~\ref{S-fig:xlinked_solvation_equilibration} of the Supporting
  Information). The `wet' equilibration and cross-linking procedures follow the
  methods outlined in our previous work.~\cite{coscia_understanding_2019}

@@

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{total_density_5wt.pdf}
  \caption{5 wt \% water, spatial density of all atoms}\label{fig:total_density_5wt}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{total_water_density_5wt.pdf}
  \caption{5 wt \% water, spatial density of water}\label{fig:total_water_density_5wt}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{total_density_10wt.pdf}
  \caption{10 wt \% water, spatial density of all atoms}\label{fig:total_density_10wt}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \vspace{-0.5cm}
  \includegraphics[width=\textwidth]{total_water_density_10wt.pdf}
  \caption{10 wt \% water, spatial density of water}\label{fig:total_water_density_10wt}
  \end{subfigure}
  \caption{Looking down at the $xy$ plane of the cross-linked LLC membranes, the lowest
  density regions of both membranes are in the pore center and in the distal tail region,
  furthest from the pore center (a \& c). Water concentration is highest in regions of
  low monomer density (b \& d).}\label{fig:total_density}
  \end{figure}

@@

  We observed transport of sodium, water and 20 other small polar solutes
  inside the membrane nanopores. First, we comment on transport of the
  membrane constituents, water and sodium, in a system absent of any additional
  solutes. We then present the general trends that we observe among 
  all other solutes studied.

@@

  % Changing does not change conclusion here.
  In the 5 wt\% water system, water molecules that spend the majority of their 
  time in the distal tail region have MSDs slightly higher than those of water molecules
  close to the pore center. The MSD of distal tail water molecules is 1.7 nm$^2$
  compared to 1.2 nm$^2$ outside the distal tails. This behavior is likely due to the
  much more crowded pore environment in the 5 wt\% versus the 10 wt\% water system. 
  Monomer head groups are concentrated about 2~\AA~closer to the pore centers in the
  5 wt\% versus the 10 wt\% water system (see Figure~\ref{fig:component_densities})  %BJC: this is the RDF figure of head groups, tails, sodium, water 
  which leaves less space for water molecules and restricts their movement among the 
  relatively immobile monomer head groups. The consequently high energetic barrier 
  to diffusion enhances the favorability of associative interactions between water
  and monomer components. The 95th percentile lifetime of hydrogen bonds between 
  water molecules located near the pore center is 2.1 ns in the 5 wt\% system compared
  to 0.5 ns in the 10 wt\% system. However, since we only saved simulation output 
  every 0.5 ns, it is likely that the true hydrogen bond lifetimes in the 10 wt\% 
  system are shorter than this value. Additionally, there are more frequent and longer
  periods of association between water and sodium ions. An average of 65\% of water 
  molecules in the pores of the 5 wt\% water system associate with a sodium ion each
  frame compared to 44\% in the 10 wt\% water system.There are about half as many water
  molecules in the 5 wt\% water system which allows a larger fraction of them to 
  associate with sodium ions. The lifetime of associations between sodium ions and 
  water is 7.5 ns and 3.0 ns in 5 and 10 wt\% water systems respectively.

@@

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\linewidth]{msd_radius_diols_10wt.pdf}
  \caption{}\label{fig:polyols_msd}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\linewidth]{polyols_rdf.pdf}
  \caption{}\label{fig:polyols_rdf}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \vspace{-0.275cm}
  \includegraphics[width=\linewidth]{multi_hbonds.pdf}
  \caption{}\label{fig:multi_hbonds}
  \end{subfigure}
  \caption{(a) The MSDs of diols, triols and sugars descend in order of their size, except
  for propylene glycol which moves exceptionally slow. The MSD of ethylene glycol 
  is in close agreement with the theoretical lines, implying that the solute 
  is subject to a similar amount of hindrance as methanol, the solute to which the
  theoretical lines were fit. (b) Glycerol, tetrose and ribose have high densities 
  close to the pore center because they have a high number of hydrophilic groups and
  are relatively large. Ethylene glycol and propylene glycol are densest close to the
  head group region. (c) The number of hydrogen bond interactions between solutes and
  monomers increases as solutes gain additional hydroxyl groups. The number of 
  hydrogen bonds made by a single solute in different locations simultaneously, n, 
  also increases with the number of hydroxyl groups. In the most extreme case, all 
  four hydroxyl groups of ribose (n = 4) are involved in a hydrogen bond interaction
  at the same time.}\label{fig:polyols}
  \end{figure}
